\documentclass[11pt]{article}
\usepackage{hyperref}
\usepackage{ulem}

\setlength{\oddsidemargin}{-0.5in}
\setlength{\evensidemargin}{\oddsidemargin}
\setlength{\textwidth}{7.0in}
\setlength{\textheight}{8.3in}
\setlength{\topmargin}{-0.7in}

%\input{macros.tex}
\begin{document}

{\Large COSC 417 \hspace{2cm} Assignment 3}
\vspace{0.5cm}

\emph{Instructions:}

(a) Due {As indicated on Blackboard.} 


(b) Submit on Blackboard.  This homework will not be graded, I will post solutions (see the syllabus for the policy regarding assignments).




		


 (c) For editing your homework. I recommend that you use Latex, see the template posted on Blackboard. 
 
         The template has an  example showing how to do the state diagram  of a finite automaton. If you have trouble with making figures in Latex, you can draw the state diagram and insert a digital photo in the .tex file (see the template how to do this).
  
	    
		   (d) If a problem has more questions, write down your answers in the same order as the order of questions. In principle, this should help you.
		   \vspace{0.5cm}
\bigskip




  Do without turning in anything the following exercises regarding regular expressions from the textbook: 1.19, 1.20, 1.21  (Some of them have solutions in the book; still try to do them before checking the solution. Some of this exercises will show up in quizzes).


  Read Examples 1.73 - 1.77  from the textbook, which use the Pumping lemma to show that certain languages are not regular. (Some of these exercises will show up in quizzes and exams).
\bigskip

Do and turn in the following exercises.

1.  Exercise 1.16  - page 86. Use the method shown in class and which is also in the textbook in Example 1.41, page 56.


\medskip

2.	Give regular expressions for each of the following languages over the alphabet $\{0,1\}$:

(a)	$\{w \in  \{0, 1\}^* \mid \mbox{ $w$ begins with a $1$ and ends with a $1$}\}$

(b)	$\{w  \in \{0, 1\}^*  \mid \mbox{ $w$ has length at least $3$ and its 2nd symbol is a $0$}\}$

(c)	$\{w  \in \{0, 1\}^*  \mid \mbox{ $w$ does not contain $01$} \}$.
\medskip


In the following $3$ exercises, use the Pumping Lemma or the closure properties of regular languages to prove that the  corresponding languages are not regular.
\medskip

3. 	$B=\{w  \in \{0, 1\}^* \mid \mbox{ the number of 0s in $w$ is two times the number of 1s in $w$}\}$.
\medskip




4. 	$A = \{0^{n!} \mid n >= 0 \}$ (Reminder: $n! = 1  \cdot 2 \cdot  \ldots \cdot  n$)
\medskip


5. 	$C = \{0^m1^n  \mid  m \not= n \}$ (Note: This is an example where the pumping lemma is difficult to use.   It is much easier to use the closure properties of regular languages.)
\medskip

{\small
(Some hints for exercises 3-5: If you use the Pumping Lemma, present how Player 1 can win the game in 3 moves. Remember that in the first move, Player 1 must choose a string that has length at least p, and p is an arbitrary number. In step 3 you must show that Player 1 has a winning move, for \emph{any} move of Player 2 in the second move. 

Exercise 3 is similar to Example 1.73 in the textbook (which we also did in class), Exercise 4 is similar to Example 1.76 in the textbook.

If you use the closure properties for Exercise 5, you need to transform (using the operations of union, intersection, complementation, concatenation, and star) the language $C$ into $L_1$, and then $L_1$ into $L_2$, and so on till you reach a language that you know already is not regular.
This implies that the initial language $C$ is also not regular.)
}
\medskip

6. Show that if $A$ is a regular language, then $A_{even} = \{x \in A \mid |x| \textrm{ is even} \}$ is also regular. \emph{Hint:} $A_{even}$ is the subset of $A$ containing all strings in $A$ with even length. You can use a closure property.
\bigskip


%The next problem is \textbf{Extra Credit}.  It is optional. If you choose to do this part, you need to do it alone (without a partner) and turn it separately in class. Write on top of it: "Assignment 3 -Extra Credit" and your name.
%\medskip

7.  Find the mistake in the following ``proof" that shows that the language $L = a^* b^*$ (formally, $L$ is the language denoted by the regular expression $a^* b^*$) is not regular (which obviously is false).
\medskip

\emph{``Proof" that $L$ is not regular}. 

Suppose $L$ is regular, and let $p$ be the pumping length for $L$. We play the Pumping Game for $L$.
\medskip

I. We choose $s = a^{p/2} b^{p/2}$ if $p$ is even. If $p$ is odd, we take $s = a^{(p+1)/2} b^{(p+1)/2}$ .
\smallskip

II. The adversary breaks $s = xyz$ so that $y=ab$.
\smallskip


III. We take $i=2$. We have won the game: $xyyz$ is not in $L$ because it has as a subword $abab$ which is not allowed for strings in $L$.
 \medskip

Since we have won the Pumping Game, it follows that $L$ is not regular.

\emph{End of ``Proof''}
 
\end{document}
 